### 2024-04-04 paper/paper.qmd
Update paper.qmd

---
title: "Attractive people, Attractive jobs "
subtitle: "Analysis of unveilling the influence of attractiveness on Employement in the United States"
author: Sky Suh
thanks: "Code and data are available at: https://github.com/sjp2023/beauty-standard.git"
date: "`r Sys.time()`"
date-format: "D MMMM YYYY"
abstract: "This paper is explored the influence of attractiveness on Employment regarding other subjects such as charactersitics, skills that individual had but by measuring the relationships with the attraictveness of physical, educational and gender for the hiring process and employments of the success. Using the data from the public dataverse source based on the paper, we found that the attractiveness impacts on employments and having higher chance to be succeed in both hriing process and careers. This paper provides how does the attractiveness of people impact on employment and the importance of the hughe biased based on the only attracitveness to predict whether attriactveness still deeply connects with the careers "
format: pdf
toc: TRUE
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
library(arrow)
library(ggplot2)
library(kableExtra)
library(knitr)

```

#### data parquet####
analysis_data<-read_parquet(here::here("data/analysis_data/analysis_data.parquet"))

# Introduction

Beauty, as the attractiveness, becomes the standard of everything in the modern society as evolving into a global conversation from body shapes, flawless skin to ideal face which influences towards individual and social-cultural perspectives. Theses standards are being as the measurements with ourselves and others. For a long time, attractiveness provides the important aspects of human interaction and shapes the judgement about people as impacting on first impression when meeting the people for the first time and may have the stereotype of what is the person like. Furthermore, these attractiveness of people influence individuals' educations, gender not only viewing the physical shapes. As a result, these types of attractiveness are creating the bias particularly in the employment. The attractive workers earn 10 to 15 percents more than average or low-looking groups [@Chand]. Even in the diverse environmental workforce of the United states, the interplay between personal attractiveness and professional success has deeply connected. The way of attractiveness could shape and lead the outcome of employment has created by the society and raised the interest among scholars, researchers, and generally regarding the impact of those connection [@Ruffle].

This paper investigates the importance of the relationship between attractiveness people and employments for both hiring process and succeed in the United States where the land of diversity and opportunity. I utilized data set from 2023 U.S. data from "University of Toronto Dataverse of Borealis" to conduct the data and analyze the connection with attractiveness people and employments regarding gender, physical and education. The estimand explored from this paper is, how does the attractiveness of people impact on employment with measured for 1 and 0 of classification 1 for being yes, 0 being no with physical, educational and gender attractiveness with the employment correlations.

Key variables found from this analysis are 2023 data of the employment rates with the physical, educational and gender attractiveness regardless individual's skills, personalities and techniques that may support for being success in the positions in U.S., as it needs to be filled a clear gap of how does attractiveness influenced the hiring process and being employed in the organization. So far, we found the attractiveness will impact for the employment regarding skills, characteristics that may be the outlines for the research and how does it impacts on composite, interview and offers but with our focused area, we were interested attractiveness and employments rate. The findings in this paper is highlighted deep connection with the beauty/attractiveness and employments in the U.S. as well as the importance points since the beauty will effect on individual's first impression and provides others' feelings and stereotype itself. By understanding this relations, it will help to acknowledge the wrong side of the society and may provide the solutions towards the researchers and policymakers or in general to provide the right facts of attractiveness and employments.

This paper is structured into the following sections: Section 2 as Data with Data Measurement, Source, and each data sets of Physical Attractiveness, Educational Attractiveness, Gender Attractiveness, and Section 3 as Model presenting the linear regression model to justify and predict the relationships between the attractiveness and employments; Section 4 shown Results of the analysis data of model linear regression graph and Section 5 providing Discussion of the implications of findings from the data set and the weakness of this paper with its next and further steps for the future study about this topic.

# Data {#sec-data}

## Data Measurement
This paper utilizes based on the R, as the statistical Programming language [@citeR], 
## Source
This dataset collected the information by survey randomly from the United States in 202 published by University of Toronto Dataverse as the part of the paper named "Attractiveness and Attainment: Status, Beauty, and Jobs in China and the United States" to get the data about the attractiveness, gender and employments to compare those variables to present the relationship with the attractiveness and employments.

## Physical Attractiveness Dataset
The physical attractiveness .vs employment dataset, represented below in @tbl-1 is shown the relationship between each variables' physical attractiveness can corresponds with the employments as higher status jobs and lower status jobs. This dataset, published by [
@USA] as filed R Data updated by 2023, collects the information from the workers in the United States with the general survey to figure out the correlation whit the attractiveness and employments and how does each other influencing, particularly, the beauty of physical that most huge part of attractiveness that people will feel or get the impression easily.  
```{r}
#| label: tbl-1
#| tbl-cap: The dataset containing physical attractiveness and the status of employment.
#| echo: false
#| warning: false

head (cleaned_data, 10) |>
  kable(
    col.names = c("higherstatusjob", "lowerstatusjob","highlyattractive", "lessattractive"),
  booktabs = TRUE
  )

```

## Educational Attractiveness Dataset
The educational attractiveness dataset, showed below in @tbl-2, presents the relationships of educational attractiveness in which the elite university in this paper case to figure out having better status of jobs/hiring come from attractive as educationally. 
```{r}
#| label: tbl-2
#| tbl-cap: The dataset containing educational attractiveness and the status of employment.
#| echo: false
#| warning: false

head (cleaned_data, 10) |>
  kable(
    col.names = c("higheruni", "middleuni","loweruni", "highlyattractive", "lessattractive"),
  booktabs = TRUE
  )

```
## Gender Attractiveness Dataset

```{r}
#| label: tbl-3
#| tbl-cap: The dataset containing gender attractiveness and the status of employment.
#| echo: false
#| warning: false

head (cleaned_data, 10) |>
  kable(
    col.names = c("male", "female", "highlyattractive", "lessattractive"),
  booktabs = TRUE
  )

```

# Model

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

\newpage

# References